---
title: "Sales Dashboard"
author: "Annabelle Petherbridge"
format:
  html:
    theme: sketchy
    toc: true
    toc-depth: 2
    number-sections: true
    version: 4
editor: visual
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Load Libraries
library(janitor) 
library(tidyverse) 
library(lubridate)

# Load Data
data <- read.csv("_site/train.csv")

# clean column names using janitor
data <- clean_names(data)

# Convert dates to Date format
data$order_date <- as.Date(data$order_date, format = "%d/%m/%Y")
data$ship_date <- as.Date(data$ship_date, format = "%d/%m/%Y")

# Calculate summary stats
total_sales <- sum(data$sales, na.rm = TRUE)
num_orders <- n_distinct(data$order_id)
avg_sales_per_order <- round(mean(data$sales, na.rm = TRUE), 2)

# Format total sales using scales package
formatted_total_sales <- scales::dollar(total_sales)

# Comma-separated number of orders
formatted_num_orders <- scales::comma(num_orders)

 # Dollar sign for avg sales
formatted_avg_sales_per_order <- scales::dollar(avg_sales_per_order) 
```

<!-- CSS for card layout -->

```{=html}
<style>
  .card-container {
    display: flex;
    justify-content: space-around;
    gap: 20px;
    flex-wrap: wrap;
    width: 100%; /* Ensure the container takes up full width */
  }

  .card {
    border: 1px solid #ccc;
    padding: 15px;
    border-radius: 8px;
    width: 45%; /* Default width for cards */
    box-shadow: 0 4px 6px rgba(0, 0, 0, 0.1);
    margin-bottom: 20px;
  }

  /* Larger cards for specific sections like Sales Over Time and Top 3 Sales Months */
  .card.sales-time {
    width: 100% !important;  /* Ensure this card takes full width */
    min-height: 500px;  /* Increase height to accommodate the graph */
    padding: 20px;
    margin-bottom: 40px; /* Extra space for large cards */
  }

  .card.top-sales {
    width: 100% !important; /* Keep this card full width if needed */
    min-height: 400px;
    padding: 20px;
  }
  
  .card.product-segment {
  width: 100% !important;
  min-height: 400px;
  padding: 20px;
  margin-bottom: 40px;
}

  @media (min-width: 768px) {
    .card {
      width: 30%; /* 3 cards in a row for larger screens */
    }

    /* Keep the larger cards at 100% width for wide plots/tables */
    .card.sales-time {
      width: 100% !important;  /* Full width for Sales Over Time on larger screens */
    }

    .card.top-sales {
      width: 48%; /* Slightly larger for top-sales */
    }
  }

  @media (max-width: 600px) {
    .card {
      width: 100%; /* Stack vertically on mobile */
    }

    .card.sales-time {
      width: 100% !important;  /* Ensure full width on mobile */
      min-height: 300px;  /* Adjust for smaller screens */
    }
  }
</style>
```
# 🛒 About the Dataset

This dashboard is built using a sales dataset from Kaggle that captures **four years of transactions** from a global superstore. Each row represents a unique order, with details spanning **customer segments**, **order and shipping dates**, **shipping methods**, **product categories**, and **sales revenue**.

The dataset enables analysis of key business insights including:

-   **Sales trends over time**

-   **Product and category profitability**

-   **Customer behavior by segment**

## 📊 Summary Stats

::: card-container
::: card
<h5 class="card-title">

💰 Total Sales

</h5>

<p class="card-text">

`r formatted_total_sales`

</p>
:::

::: card
<h5 class="card-title">

📦 Number of Orders

</h5>

<p class="card-text">

`r formatted_num_orders`

</p>
:::

::: card
<h5 class="card-title">

🧮 Avg Sales / Order

</h5>

<p class="card-text">

`r formatted_avg_sales_per_order`

</p>
:::
:::

## 🔝 Top 3 Sales Months

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Add month name and year
data <- data %>%
  mutate(
    year = year(order_date),
    month_num = month(order_date),
    month_name = month.name[month_num]
  )

# Step 1: Get total sales per calendar month per year
monthly_totals <- data %>%
  group_by(year, month_num, month_name) %>%
  summarise(monthly_sales = sum(sales, na.rm = TRUE), .groups = "drop")

# Step 2: Calculate average monthly sales across years per month
avg_month_sales <- monthly_totals %>%
  group_by(month_name, month_num) %>%
  summarise(avg_sales = mean(monthly_sales), .groups = "drop") %>%
  arrange(desc(avg_sales)) %>%
  slice_head(n = 3)

# Format the sales for display
avg_month_sales$avg_sales <- scales::dollar(round(avg_month_sales$avg_sales, 2))

# Store the top 3 months and their sales for later use
top_months <- avg_month_sales %>%
  select(month_name, avg_sales) %>%
  mutate(html = paste0("<strong>", month_name, "</strong>: ", avg_sales))
```

::: card
<h5 class="card-title">

🏆 Top 3 Sales Months (Avg Across 4 Yrs)

</h5>

<p class="card-text">

`r paste(top_months$html, collapse = "<br>")`

</p>
:::

## 📅 Sales Over Time

This chart shows total sales over time, aggregated by month

::: {.card .sales-time}
<h5 class="card-title">

Sales Over Time

</h5>

```{r sales-time, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=10, fig.show='hold'}

# Data transformation
monthly_sales <- data %>% 
  mutate(month = floor_date(order_date, "month")) %>% 
  group_by(month) %>% 
  summarise(total_sales = sum(sales, na.rm = TRUE))

# Plotting the sales over time 
ggplot(monthly_sales, aes(x = month, y = total_sales)) + 
  geom_line(color = "#2C3E50", linewidth = 1.2) + 
  labs(x = "Month", y = "Sales ($)") + theme_minimal()
```
:::

## 📊 Sales by Product Segment

This section will display a bar plot showing total sales for each product segment.

<div class="card product-segment">

<h5 class="card-title"> Sales by Product Segment</h5>

```{r, , echo=FALSE, warning=FALSE, messsage=FALSE}

# Summarize total sales by product segment
segment_sales <- data %>%
  group_by(segment) %>%
  summarise(total_sales = sum(sales, na.rm = TRUE)) %>%
  arrange(desc(total_sales))

# Create a bar plot
ggplot(segment_sales, aes(x = reorder(segment, total_sales), y = total_sales, fill = segment)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Total Sales by Product Segment",
    x = "Product Segment",
    y = "Total Sales ($)",
    fill = "Segment"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

</div>

## 📎 Appendix

This dashboard was created using data from a public Kaggle dataset: [Sales Dataset on Kaggle](https://www.kaggle.com/)
